Complete Algebraic File Synchronization

### 2015-12-06 2015/def-commands.tex
Reconciliation policy comparison

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,V)$, $edit(n,V)$ and $remove(n)$ where $n\in\setn$ and $V\in\setv$ (but $V\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC:alg} we know that a useful set of axioms
will in some cases need to distinguish between edits that result in directories ($edit(n,D)$) and
ones that result in files ($edit(n,F)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (\cite{BZ}) that extending this distinction to more commands
utlimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($D$, $F$ or $\empt$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
\emph{exact} output value of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

>>>>

Let us consider what kind of information may be encoded in the commands.
A usual set of commands, based on the most frequent tools implemented by filesystems,
might be $create(n,V)$, $edit(n,V)$ and $remove(n)$ where $n\in\setn$ and $V\in\setv$ (but $V\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between edits that result in directories ($edit(n,D)$) and
ones that result in files ($edit(n,F)$), and treat them as separate commands.
Indeed, Bill Zissimopoulos suggested (\cite{BZ}) that extending this distinction to more commands
utlimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($D$, $F$ or $\empt$) in the commands is definitely useful.
At the same time, there is never any need to encode information about the
\emph{exact} output value of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

### 2015-12-06 2015/ordering.tex
Reconciliation policy comparison

If, for example, in a component or sub-component
made up of construction pairs,
the command on the topmost node, $\cxynv$, was not before all other commands
in the component as described by the algorithm,
then any resulting sequence would necessarily contain
a $\czwmv$ where $n\descendant m$, potentially followed by commands
on unrelated paths, and then followed by $\cxynv$.
From \cref{ax_separate_commute} we know we can convert this sequence
to an equivalent sequence by swapping the unrelated commands with
$\czmv$, which sequence would therefore have the commands on
$m$ and $n$ next to each other. Then from \cref{ax_distantrel_breaks,ax_directparent_breaks}
we know the sequence would break all filesystems
as a reversed construction pair is never a valid destruction pair.

>>>>

If, for example, in a component or sub-component
made up of construction pairs,
the command on the topmost node, $\cxynv$, was not before all other commands
in the component as described by the algorithm,
then any resulting sequence would necessarily contain
a $\czwmv$ where $n\descendant m$, potentially followed by commands
on unrelated paths, and then followed by $\cxynv$.
From \cref{ax_separate_commute} we know we can convert this sequence
to an equivalent sequence by swapping the unrelated commands with
$\czwmv$, which sequence would therefore have the commands on
$m$ and $n$ next to each other. Then from \cref{ax_distantrel_breaks,ax_directparent_breaks}
we know the sequence would break all filesystems
as a reversed construction pair is never a valid destruction pair.

### 2015-12-06 2015/paper.tex
Reconciliation policy comparison

In this paper we
continue the work in \cite{NREC:alg} and add to the theoretical understanding
of synchronization based not on filesystem states, but filesystem changes or commands.

@@

@@

@@

\bibitem{NREC:alg} Ramsey, Norman and Elod Csirmaz: {\it An algebraic approach to
file synchronization...}

\bibitem{BZ} Bill Zissimopoulos, personal communication, ...

>>>>

In this paper we
continue the work in \cite{NREC} and add to the theoretical understanding
of synchronization based not on filesystem states, but filesystem changes or commands.

@@

\section{Comparisons}

We consider the reconciliation algorithm described here an improvement over
the one derved during our previous research (see \cite{NREC})
as the previous algorithm not only fails to propagate all possible commands
wherever it is possible (that is, where it does not break the filesystem),
but the current algorithm is also simpler.
This is because the previous reconciliation algorithm excludes
commands from being propagated which must be preceded by a command that conflicts.

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe between different reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the command does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\cbdaa{\nn}{\valdi}\cc\cbfaa{n}{\valf} \\
B&=\cbdaa{\nn}{\valdii}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\cbfaa{n}{\valf}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same operation in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

@@

\section{Acknowledgments}

I thank Laszlo Csirmaz for his invaluable input on previous drafts of this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.

@@

\bibitem{NREC} Ramsey, Norman and Elod Csirmaz: {\it An algebraic approach to
file synchronization...}

\bibitem{CBNR} Constantin Berzan, Norman Ramsey: Summer Scholars Technical Report
Tufts University, Medford, MA, USA, 2010
http://thirld.com/files/summerscholars\_techreport.pdf
Accessed 29 November 2015
